---
title: 'Equivalence testing'
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
# knitr::opts_knit$set(global.par=TRUE)
```

```{r klippy, echo=FALSE}
klippy::klippy(
  position = c('top', 'right'),
  tooltip_message = 'Click to copy',
  tooltip_success = 'Done'
)
```

## Getting started

### Install packages

Install required packages from CRAN and GitHub.

```{r, eval=FALSE}
install.packages(c('TOSTER', 'ecoCopula', 'mvabund', 'devtools', 'psych'))
devtools::install_github('lsjmichelle/ecopower@2b3d7fe')
```

### Load packages
```{r, message=FALSE, warning=FALSE}
library(TOSTER)
library(ecopower)
library(ecoCopula)
library(mvabund)
```

### Load data

The `crayweed` dataset was collected over a period of seven time points and
contains the counts of 20 fish species at 27 sites - 16 reference sites and
11 restored sites.

```{r}
data('crayweed', package='ecopower')
abund = crayweed$abund
X = crayweed$X
```

## TOSTER

### Species richness data

```{r}
data = data.frame(richness = rowSums(abund>0))
data$treatment = X$treatment
data$treatment = factor(data$treatment, levels=c('restored', 'reference'))
head(data)
```

### Testing for equivalence

```{r, message=FALSE}
dataTOSTtwo(
  data, deps='richness', group='treatment',
  low_eqbound=-0.8, high_eqbound=0.8,
  var_equal=TRUE, desc=TRUE, plots=TRUE
)
```

We do not find evidence of similarity. The confidence interval (CI)
does not fall between the lower and upper bounds.

### Testing for non-inferiority

```{r, message=FALSE}
dataTOSTtwo(
  data, deps='richness', group='treatment',
  low_eqbound=-0.8, high_eqbound=Inf,
  var_equal=TRUE, desc=TRUE, plots=TRUE
)
```

The result is inconclusive. To conclude that restored is not inferior to reference,
the CI should be above the lower bound.

### More examples

Our study lacks power at an effect size of 0.8.

```{r}
powerTOSTtwo(alpha=0.05, N=14, low_eqbound_d=-0.8, high_eqbound_d=0.8)
```

When designing an experiment, collecting more samples will improve the
chances of detecting an effect size of interest.

```{r}
powerTOSTtwo(alpha=0.05, statistical_power=0.8, low_eqbound_d=-0.8, high_eqbound_d=0.8)
```
<br>
`TOSTER` allows us to combine equivalence test and classical test. To do so, we require
some descriptive statistics. There are many ways to compute these statistics.
Here we use the `psych` package to display the descriptive statistics in a table.

```{r}
psych::describeBy(
  richness~treatment, mat=TRUE, digits=4, data=data
)[,c('group1', 'n', 'mean', 'sd')]
```

Then we input the values into the `TOSTtwo` function.
```{r}
TOSTtwo(
  m1=8.9091, m2=9.9375,
  sd1=2.0226, sd2=1.6919,
  n1=11, n2=16,
  low_eqbound_d=-0.8, high_eqbound_d=0.8,
  alpha=0.05, var.equal=TRUE
)
```

## ecopower

### Multivariate abundance data

```{r}
cbind(abund[1:6,2:4], head(X))
```

### Multivariate equivalence test

Fit a predictive model using the `manyglm` function.
```{r}
fit = manyglm(abund ~ time + treatment, family='negative.binomial', data=X)
```

Fit a Gaussian copula factor analytic model.
```{r}
fit.cord = cord(fit)
```

Specify the increasers and decreasers.
```{r}
increasers = c(
  'Acanthurus.nigrofuscus', 'Chromis.hypsilepis', 'Parma.microlepis',
  'Pempheris.compressa', 'Scorpis.lineolatus', 'Trachinops.taeniatus'
)

decreasers = c(
  'Aplodactylus.lophodon', 'Atypichthys.strigatus', 'Cheilodactylus.fuscus',
  'Pictilabrus.laticlavius'
)
```

Generate matrix of effect sizes based on effect size of interest.
```{r}
effect_mat = effect_alt(fit, effect_size=1.5, increasers, decreasers, term='treatment')
```

Perform a multivariate equivalence test.
```{r, eval=TRUE}
equivtest(fit.cord, coeffs=effect_mat)
```

There is no evidence of similarity between the reference and restored sites
(p-value > 0.1). This does not necessarily mean that the difference between
the treatment levels is greater than a factor of 1.5 - another explanation is
that we might not have enough information in the data to accurately estimate
the effect size.

## TL;DR (show me the code)
```{r, eval=FALSE}
# Install required packages from CRAN and GitHub.
install.packages(c('TOSTER', 'ecoCopula', 'mvabund', 'devtools', 'psych'))
devtools::install_github('lsjmichelle/ecopower@2b3d7fe')

# Load packages
library(TOSTER)
library(ecopower)
library(ecoCopula)
library(mvabund)

# Load data 
data('crayweed', package='ecopower')
abund = crayweed$abund
X = crayweed$X

# TOSTER

# Species richness data
data = data.frame(richness = rowSums(abund>0))
data$treatment = X$treatment
data$treatment = factor(data$treatment, levels=c('restored', 'reference'))
head(data)

# Testing for equivalence
dataTOSTtwo(
  data, deps='richness', group='treatment',
  low_eqbound=-0.8, high_eqbound=0.8,
  var_equal=TRUE, desc=TRUE, plots=TRUE
)

# Testing for non-inferiority
dataTOSTtwo(
  data, deps='richness', group='treatment',
  low_eqbound=-0.8, high_eqbound=Inf,
  var_equal=TRUE, desc=TRUE, plots=TRUE
)

# More examples

powerTOSTtwo(alpha=0.05, N=14, low_eqbound_d=-0.8, high_eqbound_d=0.8)

powerTOSTtwo(alpha=0.05, statistical_power=0.8, low_eqbound_d=-0.8, high_eqbound_d=0.8)

# Descriptive statistics
psych::describeBy(
  richness~treatment, mat=TRUE, digits=4, data=data
)[,c('group1', 'n', 'mean', 'sd')]

TOSTtwo(
  m1=8.9091, m2=9.9375,
  sd1=2.0226, sd2=1.6919,
  n1=11, n2=16,
  low_eqbound_d=-0.8, high_eqbound_d=0.8,
  alpha=0.05, var.equal=TRUE
)

# ecopower - Multivariate equivalence test

# Multivariate abundance data
cbind(abund[1:6,2:4], head(X))

# Fit a predictive model using the manyglm function
fit = manyglm(abund ~ time + treatment, family='negative.binomial', data=X)

# Fit a Gaussian copula factor analytic model
fit.cord = cord(fit)

# Specify the increasers and decreasers
increasers = c(
  'Acanthurus.nigrofuscus', 'Chromis.hypsilepis', 'Parma.microlepis',
  'Pempheris.compressa', 'Scorpis.lineolatus', 'Trachinops.taeniatus'
)

decreasers = c(
  'Aplodactylus.lophodon', 'Atypichthys.strigatus', 'Cheilodactylus.fuscus',
  'Pictilabrus.laticlavius'
)

# Generate matrix of effect sizes based on effect size of interest
effect_mat = effect_alt(fit, effect_size=1.5, increasers, decreasers, term='treatment')

# Perform a multivariate equivalence test
equivtest(fit.cord, coeffs=effect_mat)
```
